---
title: "Electric Vehicle Adoption in Bay Area"
output:
  html_notebook: default
  pdf_document: default
---


**Library all packages**
```{r}
library(httr)
library(jsonlite)
library(readxl)
library(dplyr)
library(ggplot2)
library(lubridate)
library(dagitty)
library(lfe)
```

**Load the air quality data from U.S. Envrionmental Protection Agency (EPA) API**
```{r}
# API key
api_key <- "dunwren65"
user_email <- "max_xu@berkeley.edu"

# Define counties and their codes
bay_area <- data.frame(County = c("Alameda", "Contra Costa", "Marin", "Napa", "San Mateo", "Santa Clara", "Solano", "Sonoma", "San Francisco"),
                       County_Code = c("001", "013", "041", "055", "081", "085", "095", "097", "075"))

# Define pollutants and their parameter codes
pollutants <- data.frame(Pollutant = c("PM2.5", "PM10", "Ozone", "NOx", "COx"),
                         Param_Code = c("88101", "81102", "44201", "42602", "42101"))

# Initialize an empty dataframe
air_quality_data <- data.frame()

# Fetch data for each county, pollutant, and year range
for (i in 1:nrow(bay_area)) {
  for (j in 1:nrow(pollutants)) {
    for (year in 2011:2023) {
       # Print progress for the current iteration
      print(paste("Processing", bay_area$County[i], pollutants$Pollutant[j], "for year", year))
      
      # Define start and end dates for each year
      bdate <- paste0(year, "0101")  # Start date (YYYYMMDD)
      edate <- paste0(year, "1231")  # End date (YYYYMMDD)
      
      # Define parameters
      params <- list(
        email = user_email,
        key = api_key,
        param = pollutants$Param_Code[j],
        bdate = bdate,
        edate = edate,
        state = "06",  # California state code
        county = bay_area$County_Code[i]
      )
      
      # API request
      response <- GET("https://aqs.epa.gov/data/api/dailyData/byCounty", query = params)
      
      # Check for successful response
      if (response$status_code == 200) {
        data <- fromJSON(content(response, as = "text", encoding = "UTF-8"))
        if (!is.null(data$Data) && nrow(as.data.frame(data$Data)) > 0) {
          # Append results to the dataframe
          county_data <- as.data.frame(data$Data)
          county_data$County <- bay_area$County[i]
          county_data$Pollutant <- pollutants$Pollutant[j]
          air_quality_data <- bind_rows(air_quality_data, county_data)
        } else {
          print(paste("No data for", bay_area$County[i], pollutants$Pollutant[j], "in year", year))
        }
      } else {
        print(paste("Error for", bay_area$County[i], pollutants$Pollutant[j], "in year", year, ":", response$status_code))
      }
    }
  }
}

# Save the combined air quality data
write.csv(air_quality_data, "air_quality_data_2011_2023.csv", row.names = FALSE)

# Preview the data
print(head(air_quality_data))

```

**Example for one year data, if the orignial one running too long**
```{r}
# API key
#api_key <- "dunwren65"
#user_email <- "max_xu@berkeley.edu"

# Define counties and their codes
#bay_area <- data.frame(County = c("Alameda", "Contra Costa", "Marin", "Napa", "San Mateo", "Santa Clara", "Solano", "Sonoma", "San Francisco"),
#                      County_Code = c("001", "013", "041", "055", "081", "085", "095", "097", "075"))

# Define pollutants and their parameter codes
#pollutants <- data.frame(Pollutant = c("PM2.5", "PM10", "Ozone", "NOx", "COx"),
#                         Param_Code = c("88101", "81102", "44201", "42602", "42101"))

# Initialize an empty dataframe
#air_quality_data <- data.frame()

# Fetch data for each county and pollutant
#for (i in 1:nrow(bay_area)) {
  #for (j in 1:nrow(pollutants)) {
    # Define parameters
   # params <- list(
    #  email = user_email,
    #  key = api_key,
    #  param = pollutants$Param_Code[j],
    #  bdate = "20230101",  
    #  edate = "20231231",  
     # state = "06",        
     # county = bay_area$County_Code[i])
    
    # API request
   # response <- GET("https://aqs.epa.gov/data/api/dailyData/byCounty", query = params)
    
    # Check for successful response
   # if (response$status_code == 200) {
    #  data <- fromJSON(content(response, as = "text", encoding = "UTF-8"))
    #  if (!is.null(data$Data) && nrow(as.data.frame(data$Data)) > 0) {
    #    # Append results to the dataframe
     #   county_data <- as.data.frame(data$Data)
     #   county_data$County <- bay_area$County[i]
     #   county_data$Pollutant <- pollutants$Pollutant[j]
     #   air_quality_data <- bind_rows(air_quality_data, county_data)
     # } else {
     #   print(paste("No data for", bay_area$County[i], pollutants$Pollutant[j]))
     # }
  #  } else {
   #   print(paste("Error for", bay_area$County[i], pollutants$Pollutant[j], ":", response$status_code))}}}

# Save the combined air quality data
#write.csv(air_quality_data, "air_quality_data.csv", row.names = FALSE)

# Preview the data
#print(air_quality_data)
```

**Load EV data that I already downloaded from California Energy Commission (CEC)**
```{r}
# Load Local Excel files path
path_to_New_ZEV_Sales <- "C:/Users/max_x/ucb/MaCSS - Final Project/New_ZEV_Sales_Last_updated_08-06-2024_ada.xlsx"
path_to_Vehicle_Population <- "C:/Users/max_x/ucb/MaCSS - Final Project/Vehicle_Population_Last_updated_04-30-2024_ada.xlsx"

# Load the data
suppressWarnings(zev_sales <- read_excel(path_to_New_ZEV_Sales, sheet = "County"))
suppressWarnings(vehicle_population <- read_excel(path_to_Vehicle_Population, sheet = "County"))

# Preview
head(zev_sales)
head(vehicle_population)
```

**Cleaning EV data**
```{r}
# Define the nine Bay Area counties
bay_area_counties <- c("Alameda", "Contra Costa", "Marin", "Napa",
                       "San Mateo", "Santa Clara", "Solano", "Sonoma", "San Francisco")

# 1. Clean and aggregate `zev_sales`
zev_sales_filtered <- zev_sales %>%
  filter(County %in% bay_area_counties) %>%  
  group_by(County, Data_Year) %>% 
  summarize(Sale_for_Zero_Emission_Vehicles = sum(`Number of Vehicles`, na.rm = TRUE)) 

# Save the cleaned dataset
write.csv(zev_sales_filtered, "zev_sales_filtered.csv", row.names = FALSE)

# 2. Clean and aggregate `vehicle_population`
vehicle_population_filtered <- vehicle_population %>%
  filter(County %in% bay_area_counties) %>%  
  group_by(County, `Data Year`, `Dashboard Fuel Type Group`) %>%  
  summarize(Total_Vehicles = sum(`Number of Vehicles`, na.rm = TRUE))  

# Save the cleaned dataset
write.csv(vehicle_population_filtered, "vehicle_population_filtered.csv", row.names = FALSE)

# Preview the cleaned datasets
print(zev_sales_filtered)
print(vehicle_population_filtered)
```

**Visualization about EV data**
```{r}
# Plot ZEV Sales Trends
ggplot(zev_sales_filtered, aes(x = Data_Year, y = Sale_for_Zero_Emission_Vehicles, color = County)) +
  geom_line(size = 1) +
  geom_point() +
  labs(title = "Trends in Zero-Emission Vehicle Sales by County (2024 only have first half year)",
       x = "Year",
       y = "Number of Zero-Emission Vehicles Sold",
       color = "County") +
  theme_minimal()
```

```{r}
# Plot Vehicle Population Trends
ggplot(vehicle_population_filtered, aes(x = `Data Year`, y = Total_Vehicles, color = County)) +
  geom_line(linewidth = 1) +  
  geom_point() +
  facet_wrap(~`Dashboard Fuel Type Group`, scales = "free_y") +
  labs(title = "Trends in Vehicle Population by County and Fuel Type",
       x = "Year",
       y = "Total Number of Vehicles",
       color = "County") +
  theme_minimal()
```

```{r}
# Filter data for a specific year (2023)
zev_sales_2023 <- zev_sales_filtered %>% filter(Data_Year == 2023)

# Bar Plot for 2023 ZEV Sales
ggplot(zev_sales_2023, aes(x = County, y = Sale_for_Zero_Emission_Vehicles, fill = County)) +
  geom_bar(stat = "identity") +
  labs(title = "Zero-Emission Vehicle Sales by County (2023)",
       x = "County",
       y = "Number of Vehicles Sold") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Filter data for a specific year (2022)
zev_sales_2022 <- zev_sales_filtered %>% filter(Data_Year == 2022)

# Bar Plot for 2022 ZEV Sales
ggplot(zev_sales_2022, aes(x = County, y = Sale_for_Zero_Emission_Vehicles, fill = County)) +
  geom_bar(stat = "identity") +
  labs(title = "Zero-Emission Vehicle Sales by County (2022)",
       x = "County",
       y = "Number of Vehicles Sold") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Combine data for 2021, 2022 and 2023
zev_sales_combined <- zev_sales_filtered %>%
  filter(Data_Year %in% c(2021, 2022, 2023))  

# Bar plot with dodge for comparison
ggplot(zev_sales_combined, aes(x = County, y = Sale_for_Zero_Emission_Vehicles, fill = factor(Data_Year))) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Zero-Emission Vehicle Sales Comparison (2021 vs 2022 vs 2023)",
       x = "County",
       y = "Number of Vehicles Sold",
       fill = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Faceted Bar Chart for EV sales
ggplot(zev_sales_filtered, aes(x = factor(`Data_Year`), y =Sale_for_Zero_Emission_Vehicles, fill = County)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ County, scales = "free_y") + 
  labs(title = "Zero-Emission Vehicle Sales by County and Year (2024 only have first half year)",
       x = "Year",
       y = "ero-Emission Vehicle Sales",
       fill = "County") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  
        strip.text = element_text(size = 10, face = "bold"))
```

**Create EV Adoption Rate**
```{r}
# Filter for Electric Vehicles
ev_population <- vehicle_population_filtered %>%
  filter(`Dashboard Fuel Type Group` %in% c("Battery Electric", "Fuel Cell (FCEV)", "Plug-in Hybrid (PHEV)")) %>%
  group_by(County, `Data Year`) %>%
  summarize(Total_EVs = sum(Total_Vehicles, na.rm = TRUE))

# Total Vehicle Population by County and Year
total_population <- vehicle_population_filtered %>%
  group_by(County, `Data Year`) %>%
  summarize(Total_Vehicles = sum(Total_Vehicles, na.rm = TRUE))

# Calculate EV Adoption Rate
ev_adoption_rate <- ev_population %>%
  left_join(total_population, by = c("County", "Data Year")) %>%
  mutate(Adoption_Rate = (Total_EVs / Total_Vehicles) * 100)

# Visualize EV Adoption Rate
ggplot(ev_adoption_rate, aes(x = `Data Year`, y = Adoption_Rate, color = County)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(title = "Electric Vehicle Adoption Rate by County",
       x = "Year",
       y = "EV Adoption Rate (%)",
       color = "County") +
  theme_minimal()
```

```{r}
# Faceted Bar Chart for EV Adoption Rates
ggplot(ev_adoption_rate, aes(x = factor(`Data Year`), y = Adoption_Rate, fill = factor(`Data Year`))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ County, scales = "free_y") +  
  labs(title = "EV Adoption Rate Trend by County and Year",
       x = "Year",
       y = "EV Adoption Rate (%)",
       fill = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        strip.text = element_text(size = 10, face = "bold"))
```

**Cleaning Air quality data**
```{r}
# Clean and refine the data
air_quality_data_cleaned <- air_quality_data %>%
  # Select relevant columns
  dplyr::select(
    date_local, County = county, Pollutant, arithmetic_mean, aqi
  ) %>%
  # Rename columns for clarity
  dplyr::rename(
    Date = date_local,
    Measurement = arithmetic_mean,
    AQI = aqi
  ) %>%
  # Convert `Date` column to Date format
  mutate(Date = as.Date(Date)) %>%
  # Remove rows with missing or invalid measurements
  filter(!is.na(Measurement) & !is.na(AQI))

# View the cleaned data
print(air_quality_data_cleaned)

# Save the cleaned dataset
write.csv(air_quality_data_cleaned, "air_quality_data_cleaned.csv", row.names = FALSE)
```

```{r}
# Remove same rows, because there are many monitors 
air_quality_data_cleaned <- air_quality_data_cleaned %>%
  distinct()

# Preview the deduplicated data
print(air_quality_data_cleaned)

# Save the deduplicated dataset
write.csv(air_quality_data_cleaned, "air_quality_data_cleaned_deduplicated.csv", row.names = FALSE)
```

```{r}
# Extract the year from the Date column
air_quality_data_cleaned <- air_quality_data_cleaned %>%
  mutate(Year = as.numeric(format(Date, "%Y")))

# Summarize the data to calculate the mean AQI for each Year and County
yearly_mean_aqi <- air_quality_data_cleaned %>%
  group_by(Year, County) %>%
  summarize(
    Mean_AQI = mean(AQI, na.rm = TRUE)  # Calculate mean AQI
  ) %>%
  ungroup()

# Preview the resulting dataset
print(yearly_mean_aqi)

# Save the summarized data to a CSV file
write.csv(yearly_mean_aqi, "yearly_mean_aqi.csv", row.names = FALSE)
```

**DAG Relationship Map for this topic**
```{r}
# DAG Relationship (just show the case, because the relationship may not be Acyclic)
dag <- dagitty("dag {EV_Adoption -> Air_Quality
                     Population_Density -> EV_Adoption
                     Population_Density -> Air_Quality
                     Industrial_Activity -> Air_Quality}")
plot(dag)
```

**t_test**
```{r}
# Subset for urban and suburban areas based on population density
urban_aqi <- air_quality_data_cleaned %>% filter(County %in% c("San Francisco"))
suburban_aqi <- air_quality_data_cleaned %>% filter(County %in% c("Alameda", "San Mateo", "Contra Costa, santa Clara"))
rural_aqi <-  air_quality_data_cleaned %>% filter(County %in% c("Solano", "Marin", "Sonoma", "Napa"))

# Perform t-test
t_test_result <- t.test(urban_aqi$AQI, suburban_aqi$AQI)
print(t_test_result)
```
```{r}
# Print Hypothesis Testing Results
cat("Hypothesis Testing:\n",
    "Null Hypothesis (H0): The mean AQI in urban areas is equal to the mean AQI in suburban areas (μ_urban = μ_suburban).\n",
    "Alternative Hypothesis (Ha): The mean AQI in urban areas is not equal to the mean AQI in suburban areas (μ_urban ≠ μ_suburban).\n\n",
    
    "Test Statistic:\n",
    "t = -18.02: The t-value indicates the difference between the group means relative to the variability in the data.\n",
    "A large absolute t-value (like -18.02) suggests a significant difference between the two groups.\n\n",
    
    "Degrees of Freedom:\n",
    "df = 34462: The degrees of freedom approximate the sample size and variability in the data.\n\n",
    
    "P-value:\n",
    "p-value < 2.2e-16: The probability of observing a difference as extreme as the one found, assuming the null hypothesis is true.\n",
    "A p-value this small indicates overwhelming evidence to reject the null hypothesis.\n\n",
    
    "Confidence Interval:\n",
    "95% Confidence Interval: [-2.382176, -1.914797]:\n",
    "The difference in mean AQI between urban and suburban areas is estimated to fall between -2.38 and -1.91 with 95% confidence.\n",
    "Since the confidence interval does not include 0, it further supports rejecting the null hypothesis.\n\n",
    
    "Sample Means:\n",
    "Mean of x (urban): 23.50.\n",
    "Mean of y (suburban): 25.65.\n",
    "Suburban areas have a higher mean AQI compared to urban areas, suggesting worse air quality in suburban areas.\n"
)
```

**Regression Model**
```{r}
print(yearly_mean_aqi)
print(ev_adoption_rate)
```
**Merge Dataset**
```{r}
# Rename columns for consistency
ev_adoption_rate <- ev_adoption_rate %>%
  rename(Year = `Data Year`)

# Merge the datasets
merged_data <- yearly_mean_aqi %>%
  inner_join(ev_adoption_rate, by = c("Year", "County"))

# Preview the merged dataset
print(merged_data)
```

**Simple Regression Model**
```{r}
# Fit a simple linear regression model
model <- lm(Mean_AQI ~ Adoption_Rate, data = merged_data)

# Display the model summary
summary(model)
```
```{r}
ggplot(merged_data, aes(x = Adoption_Rate, y = Mean_AQI)) +
  geom_point(aes(color = County)) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Relationship Between EV Adoption Rate and Mean AQI",
    x = "EV Adoption Rate (%)",
    y = "Mean AQI"
  ) +
  theme_minimal()
```

```{r}
print("Mean AQI=25.9365 − 0.9496 × Adoption Rate")
print("There is a negative relationship between EV adoption rate and air pollution, but the evidence is not strong enough for definitive conclusions.The model has a low R-squared value, suggesting other factors influence AQI significantly.")
```

**Multiple Regression Model with three separate dummy variables (Urban, Suburban, Rural)**
```{r}
# Step 1: Add a Region Category
merged_data <- merged_data %>%
  mutate(Region = case_when(
      County %in% c("San Francisco") ~ "Urban",
      County %in% c("Alameda", "San Mateo", "Contra Costa", "Santa Clara") ~ "Suburban",
      County %in% c("Solano", "Marin", "Sonoma", "Napa") ~ "Rural",
      TRUE ~ "Unknown"))

# Step 2: Add three separate dummy variables
merged_data <- merged_data %>%
  mutate(
    Urban = ifelse(County %in% c("San Francisco"), 1, 0),
    Suburban = ifelse(County %in% c("Alameda", "San Mateo", "Contra Costa", "Santa Clara"), 1, 0),
    Rural = ifelse(County %in% c("Solano", "Marin", "Sonoma", "Napa"), 1, 0)
  )

# View the updated data with new columns
print(merged_data)
```

```{r}
#Run a Multiple Regression Model
regression_model <- lm(Mean_AQI ~ Urban + Suburban + Rural + Adoption_Rate,data = merged_data)

# View the summary of the regression model
summary(regression_model)
```
```{r}
cat("Coefficients:
1. Intercept (β₀ = 25.46):
   - The predicted mean AQI for rural areas when Adoption_Rate = 0 (before EV adoption milestones).

2. Urban (β₁ = -0.9995):
   - Urban areas have, on average, a slightly lower AQI than rural areas.
   - However, the p-value (0.3442) indicates this difference is not statistically significant.

3. Suburban (β₂ = 1.558):
   - Suburban areas have a mean AQI that is 1.56 units higher than rural areas.
   - The p-value (0.0228) indicates this difference is statistically significant at the 5% level.

4. Rural (β₃ = NA):
   - The rural variable is omitted because it is the baseline category (collinearity).

5. Adoption_Rate (β₄ = -1.0863):
   - For each 1% increase in EV adoption, the mean AQI decreases by approximately 1.09 units.
   - The p-value (0.0315) indicates this effect is statistically significant at the 5% level.

Model Fit:
- Residual Standard Error: 3.366
  - Indicates the average deviation of observed AQI values from the predicted values.
  - A smaller value suggests better model accuracy.

- R-squared (R² = 0.1003):
  - About 10% of the variation in mean AQI is explained by the predictors in the model.
  - This indicates a relatively low explanatory power, suggesting other variables may influence AQI.

- Adjusted R-squared (R_adj² = 0.07603):
  - Adjusts R² for the number of predictors. Lower than R², it reflects the model's limited ability to generalize.

- F-statistic: 4.127, p-value = 0.008138
  - Tests the joint significance of all predictors.
  - The small p-value indicates the predictors are collectively significant.
")

```
```{r}
cat("Overall, EV adoption improves air quality (lower AQI), suggesting the effectiveness of promoting EV usage. While Urban areas do not show a statistically significant difference in AQI compared to Rural areas, Suburban areas exhibit significantly worse air quality, with an AQI increase of 1.558 units. This emphasizes the need to address suburban-specific contributors to air pollution.")
```

**Difference-in-Differences (DiD) analysis (pre-COVID era and the post-COVID era)**
```{r}
# Define median adoption rate for grouping
median_adoption_rate <- median(merged_data$Adoption_Rate, na.rm = TRUE)

# Add treatment indicator
merged_data <- merged_data %>%
  mutate(Treatment = ifelse(Adoption_Rate > median_adoption_rate, 1, 0))

# Preview the dataset
print(head(merged_data))
```

```{r}
# Add a time period pre-COVID era and the post-COVID era
merged_data <- merged_data %>%
  mutate(Post = ifelse(Year >= 2020, 1, 0))

# Preview the dataset
print(merged_data)
```
```{r}
# Run the DiD regression
did_model <- lm(Mean_AQI ~ Treatment * Post, data = merged_data)

# Display model summary
summary(did_model)
```
```{r}
# Aggregate data by group and year
agg_data <- merged_data %>%
  group_by(Year, Treatment) %>%
  summarize(Mean_AQI = mean(Mean_AQI, na.rm = TRUE), .groups = "drop")

# Create the plot
ggplot(agg_data, aes(x = Year, y = Mean_AQI, color = as.factor(Treatment), group = Treatment)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Difference-in-Differences Analysis: Impact of EV Adoption on AQI",
    x = "Year",
    y = "Mean AQI",
    color = "Group"
  ) +
  theme_minimal()
```

